---
title: "Exploratory data analysis"
author: "Venkata Battepati"
date: "02/03/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(psych) # Contains the winsor function and other helpful statistical tools
library(tidyverse) # Remember from last homework that dplyr is included in tidyverse
library(gridExtra)
library(corrplot)
library(patchwork)
```


Reading datasets
```{r}
positive <- read.csv("Sepsis_control_data.csv")
control <- read.csv("Sepsis_positive_data.csv")
positive$sepsis_group <- NULL
control$sepsis_group <- NULL
head(control)
head(positive)
```
Summary of datasets
```{r}
summary(positive)
summary(control)
```




Making box and violin for control age vs temp


```{r}
violin_control = ggplot(control, aes(x=age, y= temp))+
  geom_violin(outlier.color="blue")+
  labs(title="age vs temp", x="age", fill="temp")
violin_control
```

```{r}
box_control = ggplot(control, aes(x=age, y= temp))+
  geom_boxplot(outlier.color="blue")+
  labs(title="age vs temp", x="age", fill="temp")
box_control
```

Making box and violin for positive age vs temp

```{r}
violin_positive = ggplot(positive, aes(x=age, y= temp))+
  geom_violin(outlier.color="blue")+
  labs(title="age vs temp", x="age", fill="temp")
violin_positive
```

```{r}
box_positive = ggplot(positive, aes(x=age, y= temp))+
  geom_boxplot(outlier.color="blue")+
  labs(title="age vs temp", x="age", fill="temp")
box_positive
```

Scatter plot of control

```{r}
 attach(control)
omit_control = na.omit(control)
numeric_control = subset(omit_control, select=c(age, temp, weight))
plot(age, numeric_control=100000, main="Scatterplot control",
   xlab="age", ylab="Vitals", pch=19, color=numeric_control)
```

Scatter plot of positive

```{r}
attach(positive)
omit_positive = na.omit(positive)
numeric_positive = subset(omit_positive, select=c(age, temp, weight))
<<<<<<< Updated upstream
plot(age, numeric_positive=10000000, main="Scatterplot positive",
   xlab="age", ylab="Vitals", pch=19, color=numeric_positive)
```

=======
plot(numeric_positive, y=100000, main="Scatterplot positive",
   xlab="Age", ylab="Vitals", pch=19, color= "numeric_positive")
```



```{r}
library(ggplot2)
attach(positive)
numeric_positive = subset(positive, select=c(temp, weight))
p <- ggplot(numeric_positive, aes(age, numeric_positive, colour = numeric_positive)) + geom_point()
p + labs(colour = "numeric_positive")
```



>>>>>>> Stashed changes
